06_gam

### 2023-10-20 06_gam.Rmd
GAMs for ZC and PM

Load allDrifts object
```{r setup, include=FALSE}
load(allDrifts, file = "data/allDrifts.rda")
```

@@

@@

#GAM by ZC
```{r}
ZC_filt <- allDrifts %>% 
  filter(species %in% c("ZC", NA))

ZC_GAM <- gam(Wpresence>0 ~ depth + mldTemp + ttDepth + sal400,
              data = ZC_filt, family = binomial, method = "REML")
summary(ZC_GAM)
gam.check(ZC_GAM)
```

# GAM by PM
```{r}
PM_filt <- allDrifts %>% 
  filter(species %in% c("PM", NA))

PM_GAM <- gam(Wpresence>0 ~ depth + sst_mean + mldTemp + ttDepth + sal400 +
                TOL_125 + TOL_2000,
              data = PM_filt, family = binomial, method = "REML")
summary(PM_GAM)
gam.check(PM_GAM)

# Using variables from Fiedler et al. 2023 for sperm whales
PM_GAM_fiedlerVar <- gam(Wpresence>0 ~ depth + mldDepth + curl_mean + ssh_mean + sal400,
              data = PM_filt, family = binomial, method = "REML")
summary(PM_GAM_fiedlerVar)
```

>>>>

Load allDrifts object
```{r setup, include=FALSE}
load(file = "data/allDrifts.rda")
```

@@

# GAM by ZC
# Top 5 env variables from RF are:depth, sst_mean, curl_mean, mldTemp, bathyslope

```{r}
ZC_filt <- allDrifts %>% 
  filter(species %in% c("ZC", NA))

# Environ only
ZC_GAM <- gam(Wpresence>0 ~ depth + sst_mean + curl_mean + mldTemp + bathy_slope,
              data = ZC_filt, family = binomial, method = "REML")
summary(ZC_GAM)
gam.check(ZC_GAM)

ZC_GAM_env_TOL <- gam(Wpresence>0 ~ depth + sst_mean + curl_mean + mldTemp + bathy_slope +
                        TOL_125 + TOL_2000 + shipin10km + timeAfterShip,
              data = ZC_filt, family = binomial, method = "REML")
summary(ZC_GAM_env_TOL)
gam.check(ZC_GAM_env_TOL)
```

# GAM by PM
# Top 5 predictors from RF: TOL_2000, dist2slope, timeaftership, ttDepth, depth

```{r}
PM_filt <- allDrifts %>% 
  filter(species %in% c("PM", NA))

# Environ only
PM_GAM <- gam(Wpresence>0 ~ dist2slope + ttDepth + mldDepth + ssh_mean + depth,
              data = PM_filt, family = binomial, method = "REML")
summary(PM_GAM)
gam.check(PM_GAM)

PM_GAM_env_TOL <- gam(Wpresence>0 ~ dist2slope + ttDepth + depth + mldDepth + ssh_mean +
                        TOL_125 + TOL_2000 + timeAfterShip,
              data = PM_filt, family = binomial, method = "REML")
summary(PM_GAM_env_TOL)
gam.check(PM_GAM_env_TOL)
```

@@
